---
title: "R-Project.Rmd"
author: "Pritam Majumder"
date: "2024-03-30"
output: html_document
---

## 1.load the "Breast cancer" Data set and Explore first few rows and summery statistics of the dataset.

```{r}
url <-"https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/wdbc.data"
breast_cancer <- read.csv(url, header = FALSE)
head(breast_cancer)
summary(breast_cancer)
```

## 2.Data cleaning: check for missing values

```{r}
missing_values <- colSums(is.na(breast_cancer))
print(missing_values) 
```

## 3.Exploratory Data Analysis

```{r}

#relationships between these features(Radius,texture,smoothness).
##Box plots 
boxplot(V3 ~ V2, data = breast_cancer, 
        main = "Box Plot of Radius Mean for Malignant and Benign Tumors",
        xlab = "Diagnosis", ylab = "Radius Mean", col = c("red", "blue"))
par(mfrow = c(1, 2))

#Histogram of Texture Mean for Malignant Tumors.
hist(subset(breast_cancer$V4, breast_cancer$V2 == "M"), 
         main = "Histogram of Texture Mean for Malignant Tumors", xlab = "Texture Mean", ylab = "Frequency", col = "red")

#Histogram of Texture Mean for Benign Tumors.
hist(subset(breast_cancer$V4, breast_cancer$V2 == "B"), 
     main = "Histogram of Texture Mean for Benign Tumors", xlab = "Texture Mean", ylab = "Frequency", col = "blue")
par(mfrow = c(1, 2))

#Histogram of Smoothness Mean for Malignant Tumors.

hist(subset(breast_cancer$V5, breast_cancer$V2 == "M"), 
     main = "Histogram of Smoothness Mean for Malignant Tumors", xlab = "Smoothness Mean", ylab= "Frequency", col = "red")

#Histogram of Smoothness Mean for Benign Tumors.

hist(subset(breast_cancer$V5, breast_cancer$V2 == "B"), 
     main = "Histogram of Smoothness Mean for Benign Tumors", xlab = "Smoothness Mean", ylab = "Frequency", col = "blue")



#Scatter Plots.

plot(breast_cancer$V3, breast_cancer$V4, xlab = "Radius Mean", ylab ="TextureMean",    main="Scatter Plot of Radius Mean vs Texture Mean")

```


## 4.Correlation Analysis and correlation matrix.

```{r}
correlation_matrix <- cor(breast_cancer[, 3:5])  # Considering features V3 (radius_mean), V4 (texture_mean), V5 (smoothness_mean)
print(correlation_matrix)
```

# Calculate correlation matrix

```{r}

relevant_features <- breast_cancer[, c("V3", "V4", "V5")]
correlation_matrix <- cor(relevant_features)
print(correlation_matrix)
```

# Visulize the  Correlation matrix

```{r}
heatmap(correlation_matrix,
        col = colorRampPalette(c("blue", "white", "red"))(100),
        symm = TRUE,
        margins = c(7,7))
```

## 5.Malignant vs Benign Analysis Explore differences in feature distributions between malignant and benign tumors.

```{r}
boxplot(V3 ~ V2, data = breast_cancer, 
        main = "Distribution of Radius Mean by Tumor Type",
        xlab = "Diagnosis", ylab = "Radius Mean", col = c("red", "blue"))
```

## Box plot for texture_mean

```{r}

boxplot(V4 ~ V2, data = breast_cancer, 
        main = "Distribution of Texture Mean by Tumor Type",
        xlab = "Diagnosis", ylab = "Texture Mean", col = c("red", "blue"))
```

## Box plot for smoothness_mean
```{r}
boxplot(V5 ~ V2, data = breast_cancer, 
        main = "Distribution of Smoothness Mean by Tumor Type",
        xlab = "Diagnosis", ylab = "Smoothness Mean", col = c("red", "blue"))
```

## 6.Identify features that show significant variations.
# Perform t-tests for each feature

```{r}

# Perform t-tests for each feature
t_test_radius <- t.test(V3 ~ V2, data = breast_cancer)
t_test_texture <- t.test(V4 ~ V2, data = breast_cancer)
t_test_smoothness <- t.test(V5 ~ V2, data = breast_cancer)

```

## Print t-test results

```{r}

print("T-Test Results for Radius Mean:")
print(t_test_radius)
print("")
print("T-Test Results for Texture Mean:")
print(t_test_texture)
print("")
print("T-Test Results for Smoothness Mean:")
print(t_test_smoothness)

```

## Conduct two-sample t-tests to compare the means of different features between malignant and benign tumors.

```{r}

t_test_results <- lapply(breast_cancer[, c("V3", "V4", "V5")], function(feature) {
t.test(feature ~ V2, data = breast_cancer)
})

```

## Print t-test results

```{r}
for (i in 1:length(t_test_results)) {
  feature_name <- names(breast_cancer)[i + 2]  # Adjust for column index and exclude ID and Diagnosis columns
  print(paste("T-Test Results for", feature_name, ":"))
  print(t_test_results[[i]])
  cat("\n")
}

```

## 7.Perform PCA on relevant features

```{r}

relevant_features <- breast_cancer[, c("V3", "V4", "V5")]  # Subset relevant features
pca_result <- prcomp(relevant_features, scale. = TRUE)  # Perform PCA with scaling

```
## Summary of PCA

```{r}

summary(pca_result)

```
## Scatter plot
```{r}
plot(pca_result, type = "l", main = "Scatter Plot for PCA")
```

# Scatter Plot based on benign and malignant tumours.
```{r}
plot(pca_result$x[,1], pca_result$x[,2], 
     col = as.factor(breast_cancer$V2),
     xlab = "PC1", ylab = "PC2",
     main = "Scatter Plot of PC1 vs PC2 with Diagnosis Colors")
legend("topright", legend = levels(as.factor(breast_cancer$V2)), col = 1:2, pch = 1)
```

## 9.Regression Model.
```{r}
# Build logistic regression model
breast_cancer$V2<- as.factor(breast_cancer$V2)
breast_cancer[, 3:ncol(breast_cancer)] <- sapply(breast_cancer[, 3:ncol(breast_cancer)], as.numeric)
str(breast_cancer)
logistic_model <- glm(V2 ~ V3 + V4 + V5, 
                      data = breast_cancer, 
                      family = binomial(link = "logit"))
summary(logistic_model)

```

 Summary of the logistic regression model

```{r}

summary(logistic_model)
```

```{r}




## Calculate predicted probabilities

predicted_probabilities <- predict(logistic_model, newdata = breast_cancer, type = "response")



logistic_model <- glm(V2 ~ V3 + V4 + V5, 
                      data = breast_cancer, 
                      family = binomial(link = "logit"))
summary(logistic_model)
predicted_diagnosis <- predict(logistic_model, newdata = breast_cancer, type = "response")
predicted_probabilities <- predict(logistic_model, newdata = breast_cancer, type = "response")
predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)
conf_matrix <- table(breast_cancer$V2, predicted_classes)
print(conf_matrix)
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
print(paste("Accuracy:", accuracy))

# Precision
precision <- conf_matrix[2, 2] / sum(predicted_classes)
print(paste("Precision:", precision))
# Recall (Sensitivity)
recall <- conf_matrix[2, 2] / sum(breast_cancer$V2 == 1)
print(paste("Recall (Sensitivity):", recall))

# Specificity
specificity <- conf_matrix[1, 1] / sum(breast_cancer$V2 == 0)
print(paste("Specificity:", specificity))

# F1-score
f1_score <- 2 * precision * recall / (precision + recall)
print(paste("F1-score:", f1_score))

# ROC curve and AUC
library(pROC)
roc_curve <- roc(breast_cancer$V2, predicted_probabilities)
auc <- auc(roc_curve)
print(paste("AUC:", auc))
plot(roc_curve)

print(conf_matrix)



```
